### 2019-03-17 Document/Proposa.tex
nnn

\documentclass[
10pt, % Main document font size
a4paper, % Paper type, use 'letterpaper' for US Letter paper
oneside, % One page layout (no page indentation)
%twoside, % Two page layout (page indentation for binding and different headers)
headinclude,footinclude, % Extra spacing for the header and footer
BCOR5mm, % Binding correction
]{scrartcl}

\input{structure.tex}
\usepackage{cite}
\usepackage{adjustbox}
\usepackage{booktabs}
\hyphenation{Fortran hy-phen-ation} 
\captionsetup{font=footnotesize}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{tabularx, makecell}%
\renewcommand\theadfont{\normalsize\bfseries}
        \usepackage{etoolbox} %
        \AtBeginEnvironment{tabularx}{\setlist[enumerate, 1]{wide, leftmargin=*, itemsep=0pt, before=\vspace{-\dimexpr\baselineskip +2 \partopsep}, after=\vspace{-\baselineskip}}}

%----------------------------------------------------------------------------------------
%	TITLE AND AUTHOR(S)
%----------------------------------------------------------------------------------------
\title{\normalfont\spacedallcaps{}} % The article title

%\author{\spacedlowsmallcaps{Fatemeh Hadi Nezhad\textsuperscript{1}}}

%\date{2019} % An optional date to appear under the author(s)

%----------------------------------------------------------------------------------------

\begin{document}

%----------------------------------------------------------------------------------------
%	HEADERS
%----------------------------------------------------------------------------------------

\renewcommand{\sectionmark}[1]{\markright{\spacedlowsmallcaps{#1}}} % The header for all pages 
\lehead{\mbox{\llap{\small\thepage\kern1em\color{halfgray} \vline}\color{halfgray}\hspace{0.5em}\rightmark\hfil}} % The header style

\pagestyle{scrheadings} % Enable the headers specified in this block

%----------------------------------------------------------------------------------------
%	TABLE OF CONTENTS & LISTS OF FIGURES AND TABLES
%----------------------------------------------------------------------------------------

%\maketitle % Print the title/author/date block

\setcounter{tocdepth}{3} % Set the depth of the table of contents to show sections and subsections only

\tableofcontents % Print the table of contents

%----------------------------------------------------------------------------------------
%	AUTHOR AFFILIATIONS
%----------------------------------------------------------------------------------------

%\let\thefootnote\relax\footnotetext{* \textit{Department of Quantitative System Biology, University of California, Merced, United States}}

%----------------------------------------------------------------------------------------

\newpage

%----------------------------------------------------------------------------------------
%	Preparing Data
%----------------------------------------------------------------------------------------
\section{Introduction}
%------------------------------------------------
\section{Background}
% write this third 
% write a background for each aim
\subsection{TryTryp Genomes}	
\subsection{tRNA identity classifier}

%------------------------------------------------
\section{Specific Aims}
% write this first
\subsection{tRNA identity classifier}

\begin{enumerate}[noitemsep] % [noitemsep] removes whitespace between the items for a compact look
\item Predict and annotate tRNA gene models from TriTryp genomes available on? TriTrypdb, a kinetoplastid genome database. 
\item Create a tRNA identity classifier based on Bayesian Networks that accepts phylogenetically structured data as input and outputs posterior probabilities of functional identities for query tRNA sequences.
\item Investigate anticodon shift/functional conversion events in tRNA genes of TryTrypDB, fly, yeast, worm, etc.

\end{enumerate} 
\subsection{reconstructing ancestral rearrangements of tRNA gene clusters in eukaryotic genomes}
Develop algorithm(s) to reconstruct ancestral rearrangements of tRNA gene clusters in eukaryotic genomes, including functional conversions and genic sequence conversion events, and apply these to the above-named eukaryotic datasets to discover functionally converting genes in these datasets and predict boundaries of gene conversion events occurring in them. (I need to share with you the manuscript from Julie’s thesis on what we discovered in Drosophila, for you to fully see the significance of this proposed project). 
\subsection{Developing a machine learning framework to model the evolution of tRNA genes on an input phylogenetic tree}
Develop a machine learning framework to model the evolution of (either or both: {consensus structure, structure-function map}) tRNA genes on an input phylogenetic tree, and use this framework to improve alignment and gene-finding of evolutionarily diverse tRNA gene-sets.
%------------------------------------------------
\section{Approach}
\subsection{tRNA identity classifier}
% for each aim write an approach

%------------------------------------------------
\section{Work to Date}
\subsection{tRNA identity classifier}
% predicted tRNA models using Aragorn and tRNAScan-SE2. 
% tRNA gene Annotation: 
% 1. Summary of predicted TryTryp tRNA genes 
% 2. Integrating the gene model results: 
%     a. Assigning functional identity to the predicted genes.
%     b. Assigning coordinates to the genes  
%	  c. vidualizing the distribution of genes models predicted for each genome 
% Create a tRNA identity classifier
% 1. built a pipeine for the alignment of tRNA models.  
% 2. built a CM profiles for each model using leave one out method. 
% Investigate anticodon shift/functional conversion events in tRNA genes of TryTrypDB
% ? how do we use bubble plots and KLD and ID logo ? study the difference of HOMOC and TriTryp
%------------------------------------------------
\section{Timeline and Milestones}
%------------------------------------------------
\section{Feasibility and Potential Pitfalls}
%------------------------------------------------
\section{Significance}
\subsection{tRNA identity classifier}
% write the significance of each aim
%----------------------------------------------------------------------------------------
%	BIBLIOGRAPHYyour own dat
%----------------------------------------------------------------------------------------
\newpage
\renewcommand{\refname}{\spacedlowsmallcaps{References}} % For modifying the bibliography heading

\bibliographystyle{unsrt}

\bibliography{sample.bib} % The file containing the bibliography

%----------------------------------------------------------------------------------------

\end{document}